---
output: html_document
---

## Analysis of mushroom growth rates

```{r,echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
library(broom)
library(modelr)
library(GGally)
df <- read_csv("../data-course/Repository/Data/mushroom_growth.csv")
```


This data set was collected to determine thefastest conditions for growing Oyster Mushrooms for a farmers' market.

The variables measured were:

```{r,echo=FALSE,message=FALSE,warning=FALSE}
names(df)
```

Let's take a quick look at our data:

```{r,echo=FALSE,message=FALSE,warning=FALSE,fig.height=3,fig.width=3}
ggpairs(df)
```

Now, the analysis! First we want to fit a couple different models:

```{r,echo=FALSE,message=FALSE,warning=FALSE}
mod1 <- glm(data=df,
            formula = GrowthRate ~ Light + Humidity)

mod2 <- glm(data=df,
            formula = GrowthRate ~ Light * Nitrogen * Species)
```

Our model formulae are:

```{r,echo=FALSE}
paste0("MOD1: ",mod1$call[2])
paste0("MOD2: ",mod2$call[2])
```

Which one is better?

```{r,echo=FALSE}
rmse_mod1 <- residuals(mod1)^2 %>% mean() %>% sqrt()
rmse_mod2 <- residuals(mod2)^2 %>% mean() %>% sqrt()

paste0("The root MSE for MOD1: ",round(rmse_mod1,3))
paste0("The root MSE for MOD2: ",round(rmse_mod2,3))
```
Our first model appears to do a better, and simpler job of explaining growth rate, so we will stick with that for predictions.

```{r,echo=FALSE,fig.width=4,fig.height=4}
add_predictions(df, mod1) %>% 
  ggplot(aes(x=Light,color=Humidity)) +
  geom_point(aes(y=GrowthRate),alpha=.25) +
  geom_point(aes(y=pred),size=5) +
  theme_bw() +
  labs(caption = "Large points are model predictions")
```

Okay, so it appears that for both species, increasing light amount will bring faster growth.

```{r}
kableExtra::kable(tidy(mod1))
```

