---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
```


## This is an example final project


I'm going to look at the good ol' mpg dataset built into the tidyverse package suite.

The structure of the dataset is as follows:
```{r}
options(scipen = 999)
library(tidyverse)
mpg %>% head()
theme_set(theme_minimal())
```

One or two of the variables aresn't easy to distinguish via their encoded values. For example, the drv (drive) type lists "f" "r" and "4", but these really refer to "front" "rear" and "4-wheel" drive systems on these cars. I can change those to be more readable like this ...

```{r, echo=TRUE}
mpg <- mpg %>% 
  mutate(drv = case_when(drv == "f" ~ "front",
                         drv == "r" ~ "rear",
                         drv == "4" ~ "four-wheel"))
mpg %>% head()
```

___

### Plotting

I can make a few plots now just beacuse that is what I do ...

A plot of city miles per gallon as a function of drive type:
```{r}
mpg %>% 
  ggplot(aes(x=drv,y=cty,fill=drv)) +
  geom_boxplot() +
  labs(fill="Drive type",
       x="Drive type",
       y="City miles-per-gallon") +
  theme(legend.position = c(.8,.8),
        panel.grid = element_blank())

```

Okay, well, are those actually different from each other?

```{r}
mpg %>% 
  aov(formula = cty ~ drv, data=.) %>% 
  report::report()

```
